SFJ: An implementation of Semantic Featherweight Java

### 2020-04-27 conference/camera ready/paper.tex
splitting the algorithm

% Algorithm typesetting package
\usepackage[plain, shortend, linesnumbered]{algorithm2e}
\SetAlFnt{\sffamily \small}
\SetKwRepeat{Do}{do}{while}
\SetKwProg{Fn}{Function}{:}{end}
\SetKwInOut{Input}{Input}\SetKwInOut{Output}{Output}

%% For section renaming in autoref
\addto\extrasenglish{%
  \renewcommand{\sectionautorefname}{\S}%
  \renewcommand{\subsectionautorefname}{\S}%
  \renewcommand{\algorithmautorefname}{Algo.}%
  %\renewcommand{\exampleautorefname}{Ex.}%
  \renewcommand{\equationautorefname}{Map}%
}

% Non standard
%\usepackage{varacca}
\input{semmacros}

@@

Some comments on \autoref{alg:subtyping} follow.
In function $generateRelation$ we iterate over the set of classes in an SFJ program.
If the class currently being processed contains types in its fields or methods not present in the subtyping relation (lines 5, 30, 42), then we add the current class to the list of \emph{unprocessed} classes (line 6) so we can process its fields and methods first and the class itself later after having all required type information.
The set of unprocessed classes will then be inspected again in a recursive call (line 10).
The next two functions of the algorithm, \emph{addClass} and \emph{checkSuperSet}, check subtyping for the current $class$ being processed and update $relation$, which is a mapping from a type to its subtypes and originally only consists of entries from \autoref{eq:initialSubtype}.
In function $addClass(class)$ we check if the type $class$ is a subtype of an existing type in $relation$ (lines 15-18), as well as the opposite, meaning if $class$ is a supertype of an existing type in $relation$ (line 19).
In order to do so $checkSuperSet$ checks all fields (lines 28-39) and all methods (lines 40-51) in $class$ and compares them with an \emph{existingClass} in $relation$.
If a subtyping relation is established, then it is added to $relation$ (line 53).
Finally, upon returning from $checkSuperSet$, we also add $class$ to its own relation (line 21) to satisfy reflexivity and to $Universe$ (line 22), which is a supertype of all types.

\begin{algorithm}
    \caption{Semantic Subtyping for SFJ Classes}
    \label{alg:subtyping}

    \DontPrintSemicolon
    \Input{$classes$, the set of classes in an SFJ program.\\
    $relation$, the mapping of types to the set of subtypes, initially Map \ref{eq:initialSubtype}. \;
    }
    \Begin{
        \Fn{generateRelation(classes: List<Class>)}{
            $unprocessed: List<Class>\ \longleftarrow []$

            \For{class in classes}{
                \If{addClass(class) = $False$}{
                    $unprocessed.add(class)$
                }
            }
            \If{$untyped \neq []$}{
                $generateRelation(unprocessed)$
            }
        }
        \;
        \Fn{addClass(class: Class) $\rightarrow$ boolean}{
            \For{existing class type in relation}{
                \If{checkSuperSet(class, existingClass) = False}{
                    \KwRet $False$\;
                }
                $checkSuperSet(existingClass, class)$\;
            }
            $relation[class].add(class)$ \;
            $relation[Universe].add(class)$ \;
            \KwRet $True$ \;
        }
        \;
        \Fn{checkSuperSet(class: Class, other: Class) $\rightarrow$ boolean}{
            $flag \longleftarrow True$ \;
            \For{field in class}{
                \If{field contains type not in relation}{
                    \KwRet $False$
                }
                \uIf{other does not contain field}{
                    $flag \longleftarrow False$
                }
                \Else{
                    \If{other.field.types does not fully contain field.types}{
                        $flag \longleftarrow False$
                    }
                }
            }

@@

            \If{$flag = True$}{
                $relation[other].add$
            }
        }
    }
\end{algorithm}

>>>>

% Algorithm typesetting package
\usepackage[plain, shortend, linesnumbered, ruled]{algorithm2e}
\SetAlFnt{\sffamily \small}
\SetKwRepeat{Do}{do}{while}
\SetKwProg{Fn}{Function}{:}{end}
\SetKwInOut{Input}{Input}\SetKwInOut{Output}{Output}

%% For section renaming in autoref
\addto\extrasenglish{%
  \renewcommand{\sectionautorefname}{\S}%
  \renewcommand{\subsectionautorefname}{\S}%
  \renewcommand{\algorithmautorefname}{Algo.}%
  %\newcommand{\exampleautorefname}{Ex.}%
  \def\exampleautorefname{Ex.}% 
  \renewcommand{\equationautorefname}{Map}%
}

% Non standard
\input{semmacros}

@@

\begin{algorithm}[t]
    \caption{Semantic Subtyping for SFJ---\emph{generateRelation}}
    \label{alg:subtyping}
    \SetKwBlock{Begin}{begin}{}
    \DontPrintSemicolon
    \Input{$classes$, the set of classes in an SFJ program.\\
    $relation$, the mapping of types to the set of subtypes, initially Map \ref{eq:initialSubtype}. \;
    }
    \Begin{
        \Fn{generateRelation(classes: List<Class>)}{
            $unprocessed: List<Class>\ \longleftarrow []$

            \For{class in classes}{
                \If{addClass(class) = $False$}{
                    $unprocessed.add(class)$
                }
            }
            \If{$untyped \neq []$}{
                $generateRelation(unprocessed)$
            }
        }
        \;
    }
\end{algorithm}
%%
%%
\setcounter{algocf}{1}
\begin{algorithm}
    \caption{Semantic Subtyping for SFJ---\emph{addClass} and \emph{checkSuperSet}}

    \DontPrintSemicolon
    \setcounter{AlgoLine}{12}
    \SetKwBlock{Begin}{}{end}
    \Begin{
        \Fn{addClass(class: Class) $\rightarrow$ boolean}{
            \For{existing class type in relation}{
                \If{checkSuperSet(class, existingClass) = False}{
                    \KwRet $False$\;
                }
                $checkSuperSet(existingClass, class)$\;
            }
            $relation[class].add(class)$ \;
            $relation[Universe].add(class)$ \;
            \KwRet $True$ \;
        }
        \;
        \Fn{checkSuperSet(class: Class, other: Class) $\rightarrow$ boolean}{
            $flag \longleftarrow True$ \;
            \For{field in class}{
                \If{field contains type not in relation}{
                    \KwRet $False$
                }
                \uIf{other does not contain field}{
                    $flag \longleftarrow False$
                }
                \Else{
                    \If{other.field.types does not fully contain field.types}{
                        $flag \longleftarrow False$
                    }
                }
            }

@@

            \If{$flag = True$}{
                $relation[other].add$
            }
        }
    }
\end{algorithm}
Some comments on \autoref{alg:subtyping} follow.
In function $generateRelation$ we iterate over the set of classes in an SFJ program.
If the class currently being processed contains types in its fields or methods not present in the subtyping relation (lines 5 and 30, 42 in the continuation of the algorithm in the next page), then we add the current class to the list of \emph{unprocessed} classes (line 6) so we can process its fields and methods first and the class itself later after having all required type information.
The set of unprocessed classes will then be inspected again in a recursive call (line 10).

The next two functions of the algorithm, \emph{addClass} and \emph{checkSuperSet} given in the next page, check subtyping for the current $class$ being processed and update $relation$, which is a mapping from a type to its subtypes and originally only consists of entries from \autoref{eq:initialSubtype}.
In function $addClass(class)$ we check if the type $class$ is a subtype of an existing type in $relation$ (lines 15-18), as well as the opposite, meaning if $class$ is a supertype of an existing type in $relation$ (line 19).
In order to do so $checkSuperSet$ checks all fields (lines 28-39) and all methods (lines 40-51) in $class$ and compares them with an \emph{existingClass} in $relation$.
If a subtyping relation is established, then it is added to $relation$ (line 53).
Finally, upon returning from $checkSuperSet$, we also add $class$ to its own relation (line 21) to satisfy reflexivity and to $Universe$ (line 22), which is a supertype of all types.
